---
title: Project 1
author: "Medha Bahukhandi"
date: "10/10/2020"
---





## PROJECT ONE : REPRODUCIBLE RESEARCH




**MAking Global Options:**

```{r setoptions, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "asis", cache = TRUE)

```






**1. Loading the data into R **


```{r load}

actdata <- read.csv("./repdata_data_activity/activity.csv", header = TRUE)
  
actdata$date<- strptime(actdata$date, "%Y-%m-%d")


```





**2. Finding the mean number of steps taken per day**



```{r plot1}

#Loading the necessary packages:
library(dplyr)
library(RColorBrewer)

#Choosing the colors:
cols<- brewer.pal(5, "Set2")


#Grouping data by date and finding sum:
 actsum<-actdata%>%
  group_by(date)%>%
  summarise(steps = sum(steps))
 
 
 #Making the histogram
 hist(actsum$steps, main = "Total Number of steps each day", xlab = "Steps", ylab = "Frequency",
     col = cols)

```





**3. Mean and Median of Steps taken per day**



```{r mean}

Mean<-  round(mean(actsum$steps, na.rm = TRUE), digits = 2)
Mean


```
```{r}
Median<- median(actsum$steps, na.rm = TRUE)
Median
```




**4. Average Daily Pattern**



```{r}

library(ggplot2)


#Loading the font:
windowsFonts(Times=windowsFont("TT Times New Roman"))

#Modifications:
cols1<- element_text(colour = "red", hjust = 0.5, family = "Times")


#Averaging steps by grouping by intervals:
msteps<- actdata%>%
  group_by(interval)%>%
  summarise(steps = mean(steps, na.rm=TRUE))


#Making the plot:
g<- ggplot(msteps, aes(interval, steps))
g+ geom_line(color= "blue")+ theme_bw()+ theme(plot.title = cols1, axis.title = cols1)+ 
  labs(title = "Average Daily Activity pattern", x = "Intervals", y = "Steps")


```



**5. Five Minute interval containing the maximum number of steps**



```{r}
msteps[which.max(msteps$steps),]$interval
```



**6. The total number of missing values in the data set:**



```{r missing}
nomissing<- sum(is.na(actdata))
nomissing
```




**7. Filling in the missing values**



```{r Impute}


meanie<- function(interval){
  msteps[msteps$interval==interval,]$steps
}

actdata1 <- actdata

for(i in 1:nrow(actdata1)){
  if (is.na(actdata1[i,]$steps)){
  actdata1[i,]$steps <- meanie(actdata1[i,]$interval)
}
}
```




**8. New histogram**



```{r}
actdata1$date<- as.Date(actdata1$date, "%Y-%m-%d")

pal<- brewer.pal(5, "Set1")

newhist<- actdata1%>%
  group_by(date)%>%
  summarise(steps= sum(steps))

hist(newhist$steps, col = pal, xlab = "Number of Steps", main = "Total Steps at different dates",
     ylab = "Frequency")
```




**9. Making factors of Weekdays and weekends **



```{r}


library(lubridate)
actdata1$days <- weekdays(actdata1$date)
 
for(i in 1:nrow(actdata1)){
  if (actdata1[i,]$days%in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")){
    actdata1[i,]$days <- "Weekday"
  }
  else{
    actdata1[i,]$days <- "Weekend"
  }
}

```




###Panel plot



```{r}
actdata1$steps<- round(actdata1$steps, digits = 2)

wsteps<- actdata1%>%
  group_by(days,interval)%>%
  summarise(steps = mean(steps, na.rm = TRUE))
  


library(ggplot2)
pal1<- element_text(hjust = 0.5, colour = "red")

g<- ggplot(wsteps, aes(interval, steps))
g+ facet_grid(days ~ .) + theme_classic()+ geom_line(color = "blue") + theme(axis.title = pal1, plot.title = pal1) +
  labs(title = "Average number of steps taken per 5-minute interval across weekdays and weekends", x = "Interval", y = "Total steps") 
```







